---
title: "analysing_long_transcripts_saturation"
output: html_document
date: "2023-08-01"
editor_options: 
  chunk_output_type: console
---

Finding saturation levels, plotting read lengths for long transcripts which were selected earlier. 

Loading Libraries
```{r warning=FALSE, message=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(GenomicFeatures)
  library(rtracklayer)
  library(dplyr)
  library(Rsamtools)
  library(glmnet)
  library(ggpubr)
  library(visreg)
  library(DESeq2)
  library(apeglm)
})
```

Loading data
```{r warning=FALSE, message=FALSE}
hdeg_normal <- read.csv("/home/bhavika/Desktop/Nanograd/Data_2/heavily_degraded.csv")
hdeg_nano <- read.csv("/home/bhavika/Desktop/Nanograd/Data_2/primary_align_nano_hdeg_pass1.csv")

udeg_normal <- read.csv("/home/bhavika/Desktop/Nanograd/Data_2/undegraded.csv")
udeg_nano <- read.csv("/home/bhavika/Desktop/Nanograd/Data_2/primary_align_nano_udeg_pass1.csv")
```

Selecting required column
```{r warning=FALSE, message=FALSE}
test_1hno <- hdeg_normal %>% 
  dplyr::select(Read, Transcript, ReadLen, end_without_sc)
test_1uno <- udeg_normal %>% 
  dplyr::select(Read, Transcript, ReadLen, end_without_sc)

test_1hna <- hdeg_nano %>% 
  dplyr::select(Read, Transcript, ReadLen, end_without_sc)
test_1una <- udeg_nano %>% 
  dplyr::select(Read, Transcript, ReadLen, end_without_sc)
```

Filtering transcripts
```{r warning=FALSE, message=FALSE}
test_filter_hno <- test_1hno %>% 
  filter(Transcript=="ENST00000344063.7")
test_filter_uno <- test_1uno %>% 
  filter(Transcript=="ENST00000344063.7")

test_filter_hna <- test_1hna %>% 
  filter(Transcript=="ENST00000344063.7")
test_filter_una <- test_1una %>% 
  filter(Transcript=="ENST00000344063.7")
```

distribution plot of the 3' ends - histogram
```{r warning=FALSE, message=FALSE}
ggplot(test_filter_hno, aes(x=end_without_sc)) + geom_histogram(binwidth = 1, col="blue") + labs(x="3' end without sc") + ggtitle("Hdeg-ENST00000344063.7") + theme_bw()

ggplot(test_filter_uno, aes(x=end_without_sc)) + geom_histogram(binwidth = 1, col="blue") + labs(x="3' end without sc") + ggtitle("Udeg-ENST00000344063.7") + theme_bw()

ggplot(test_filter_hna, aes(x=end_without_sc)) + geom_histogram(binwidth = 1, col="blue") + labs(x="3' end without sc") + ggtitle("Hdeg with Nano-ENST00000361575.4") + theme_bw()

ggplot(test_filter_una, aes(x=end_without_sc)) + geom_histogram(binwidth = 1, col="blue") + labs(x="3' end without sc") + ggtitle("Udeg with Nano-ENST00000361575.4") + theme_bw()
```

Ordering the 3' end column and calculating the frequency of each 3' end
```{r warning=FALSE, message=FALSE}
test_order_hno <- test_filter_hno[order(test_filter_hno$end_without_sc, decreasing = TRUE),]
test_order_uno <- test_filter_uno[order(test_filter_uno$end_without_sc, decreasing = TRUE),]
test_order_hna <- test_filter_hna[order(test_filter_hna$end_without_sc, decreasing = TRUE),]
test_order_una <- test_filter_una[order(test_filter_una$end_without_sc, decreasing = TRUE),]

test_2hno <- test_order_hno %>% 
  group_by(end_without_sc) %>% 
  mutate(number=n())
test_2uno <- test_order_uno %>% 
  group_by(end_without_sc) %>% 
  mutate(number=n())

test_2hna <- test_order_hna %>% 
  group_by(end_without_sc) %>% 
  mutate(number=n())
test_2una <- test_order_una %>% 
  group_by(end_without_sc) %>% 
  mutate(number=n())
```

Taking the cumulative sum of the frequency 
Undegraded- without nanocount
```{r warning=FALSE, message=FALSE}
test_2uno$cum <- cumsum(test_2uno$number)
p <- function(x,n){
  len.diff <- n - length(x)
  c(rep(NA, len.diff), x)
}
test_2uno$dif <- p(diff(test_2uno$cum, lag=20), length(test_2uno$cum))
test_2uno$dif[is.na(test_2uno$dif)] <- 0

test_2uno$saturation <- test_2uno$end_without_sc[which.max(test_2uno$dif)] + 20
```

Undegraded- with nanocount
```{r warning=FALSE, message=FALSE}
test_2una$cum <- cumsum(test_2una$number)
p <- function(x,n){
  len.diff <- n - length(x)
  c(rep(NA, len.diff), x)
}
test_2una$dif <- p(diff(test_2una$cum, lag=20), length(test_2una$cum))
test_2una$dif[is.na(test_2una$dif)] <- 0

test_2una$saturation <- test_2una$end_without_sc[which.max(test_2una$dif)] + 20
```

Heavy degraded- without Nanocount
```{r warning=FALSE, message=FALSE}
test_2hno$cum <- cumsum(test_2hno$number)
p <- function(x,n){
  len.diff <- n - length(x)
  c(rep(NA, len.diff), x)
}
test_2hno$dif <- p(diff(test_2hno$cum, lag=20), length(test_2hno$cum))
test_2hno$dif[is.na(test_2hno$dif)] <- 0

test_2hno$saturation <- test_2hno$end_without_sc[which.max(test_2hno$dif)] + 20
```

Heavy degraded- with Nanocount
```{r warning=FALSE, message=FALSE}
test_2hna$cum <- cumsum(test_2hna$number)
p <- function(x,n){
  len.diff <- n - length(x)
  c(rep(NA, len.diff), x)
}
test_2hna$dif <- p(diff(test_2hna$cum, lag=20), length(test_2hna$cum))
test_2hna$dif[is.na(test_2hna$dif)] <- 0

test_2hna$saturation <- test_2hna$end_without_sc[which.max(test_2hna$dif)] + 20
```

