---
title: "finding_region_95%_reads"
output: html_document
date: "2023-04-28"
editor_options: 
  chunk_output_type: console
---

This script is about finding where 95% of the reads lie in all three conditions on the subset of the data. 

Loading Libraries
```{r warning=FALSE, message=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(GenomicFeatures)
  library(rtracklayer)
  library(dplyr)
  library(Rsamtools)
  library(glmnet)
  library(ggpubr)
  library(visreg)
  library(DESeq2)
  library(apeglm)
})
```

Loading the subset data- JCSMR Machine
```{r warning=FALSE, message=FALSE}
undegraded <- read.csv("/home/bhavika/Desktop/Nanograd/Data/csv_files/subset_un.csv")
degraded_heavy <- read.csv("/home/bhavika/Desktop/Nanograd/Data/csv_files/subset_heavy.csv")
mildly_degraded <- read.csv("/home/bhavika/Desktop/Nanograd/Data/csv_files/subset_md.csv")
```

creating a new column in all the datasets for condition
```{r warning=FALSE, message=FALSE}
degraded_heavy$condition <- "heavily deg"
mildly_degraded$condition <- "mildly deg"
undegraded$condition <- "undegraded"
```

Merging all the datasets into one
```{r warning=FALSE, message=FALSE}
a <- merge(x=undegraded, y=degraded_heavy, all=T)
combined <- merge(x=a, y=mildly_degraded, all=T)
```

'combined' dataset has all three conditions.  

fitting a model of map len vs del len
```{r warning=FALSE, message=FALSE}
model_1 <- lm(DelLen ~ MapLen, data = combined)

ggplot(combined, aes(x=MapLen, y=DelLen)) + geom_point() + labs(x="Map Len", y="Del Len") + theme_bw() + stat_ellipse(geom = "polygon", alpha=0.2, level = 0.95, colour="red") + geom_smooth(method = "lm")  + ylim(0,400) + xlim(0,3000) 
```

Plotting separate linear models from all three data frames of map len vs del len
```{r warning=FALSE, message=FALSE}
model_1 <- lm(DelLen ~ MapLen, data = undegraded)
model_2 <- lm(DelLen ~ MapLen, data = degraded_heavy)
model_3 <- lm(DelLen ~ MapLen, data = mildly_degraded)

ggplot() +
  geom_smooth(aes(x=MapLen, y=DelLen), data = undegraded, method = "lm", se=FALSE, color="red") +
  geom_smooth(aes(x=MapLen, y=DelLen), data = degraded_heavy, method = "lm", se=FALSE, color="blue") +
  geom_smooth(aes(x=MapLen, y=DelLen), data = mildly_degraded, method = "lm", se=FALSE, color="green") +
  geom_point(aes(x=MapLen, y=DelLen), data = undegraded, colour = "red") +
  geom_point(aes(x=MapLen, y=DelLen), data = degraded_heavy, colour = "blue") +
  geom_point(aes(x=MapLen, y=DelLen), data = mildly_degraded, colour = "green") + xlim(0,3500) + ylim(0,400)

```

Using Visreg package

Fitting the model from combined data set using visreg package 
1. Map len and del len
```{r warning=FALSE, message=FALSE}
fit_1 <- lm(DelLen ~ MapLen + condition, data = combined)
fit_2 <- lm(DelLen ~ MapLen, data = combined)

visreg(fit_1, xvar = "MapLen", by="condition", overlay=T)

anova(fit_1, fit_2, test="LRT")
```

2. Read len and del len
```{r warning=FALSE, message=FALSE}
fit_3 <- lm(DelLen ~ ReadLen + condition, data = combined)
fit_4 <- lm(DelLen ~ ReadLen, data = combined)

visreg(fit_3, xvar = "ReadLen", by="condition", overlay=T)

anova(fit_3, fit_4, test="LRT")
```

4. map len and insert len
```{r warning=FALSE, message=FALSE}
fit_1 <- lm(InsertLen ~ MapLen + condition, data = combined)
fit_2 <- lm(InsertLen ~ MapLen, data = combined)

visreg(fit_1, xvar = "MapLen", by="condition", overlay=T)

anova(fit_1, fit_2, test="LRT")
```

5. max del len and max map len- map len above 1000 and above 400
filtering map len and del len according to their max range
```{r warning=FALSE, message=FALSE}
test_1 <- combined %>% 
  filter(MapLen > 1000) %>% 
  filter(DelLen > 400)
```

```{r warning=FALSE, message=FALSE}
fit_3 <- lm(DelLen ~ MapLen + condition, data = test_1)
fit_4 <- lm(DelLen ~ MapLen, data = test_1)

visreg(fit_3, xvar = "MapLen", by="condition", overlay=T)

anova(fit_3, fit_4, test="LRT")
```

