---
title: "rat_cerebellum_edgeR"
output: html_document
date: "2023-10-09"
editor_options: 
  chunk_output_type: console
---

Running edgeR first on unfiltered undeg and hdeg data

loading libraries
```{r warning=FALSE, message=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(GenomicFeatures)
  library(rtracklayer)
  library(dplyr)
  library(Rsamtools)
  library(glmnet)
  library(ggpubr)
  library(visreg)
  library(DESeq2)
  library(apeglm)
  library(edgeR)
  library(coriell)
})
```

loading data
```{r warning=FALSE, message=FALSE}
rat_low1 <- set_1
rat_low2 <- set_1
rat_high1 <- set_1
rat_high2 <- set_1
```

adding the name of the condition column in all the datasets
```{r warning=FALSE, message=FALSE}
rat_low1$condition <- "low_rin_1"
rat_low2$condition <- "low_rin_2"
rat_high1$condition <- "high_rin_1"
rat_high2$condition <- "high_rin_2"
```

selecting the required columns
```{r warning=FALSE, message=FALSE}
rh_1 <- rat_high1 %>% 
  dplyr::select(Transcript, condition)
rl_1 <- rat_low1 %>% 
  dplyr::select(Transcript, condition)
rh_2 <- rat_high2 %>% 
  dplyr::select(Transcript, condition)
rl_2 <- rat_low2 %>% 
  dplyr::select(Transcript, condition)
```

Grouping the data by transcripts and then find the number of reads of each transcript and reading the column condition as to calculate number of reads using summarise command
```{r warning=FALSE, message=FALSE}
rh_grp1 <- rh_1 %>% 
  dplyr::group_by(Transcript) %>% 
  summarise(nreads=n())
rh_grp1$condition <- "high_rin_1"

rl_grp1 <- rl_1 %>% 
  dplyr::group_by(Transcript) %>% 
  summarise(nreads=n())
rl_grp1$condition <- "low_rin_1"

rh_grp2 <- rh_2 %>% 
  dplyr::group_by(Transcript) %>% 
  summarise(nreads=n())
rh_grp2$condition <- "high_rin_2"

rl_grp2 <- rl_2 %>% 
  dplyr::group_by(Transcript) %>% 
  summarise(nreads=n())
rl_grp2$condition <- "low_rin_2"
```

Merging the data to get one dataset for all the number of reads of each transcript in each condition
```{r warning=FALSE, message=FALSE}
deg_1 <- merge(x=rh_grp1, y=rl_grp1, all=T)
deg_2 <- merge(x=rh_grp2, y=rl_grp2, all=T)

data_hu <- merge(x=deg_1, y=deg_2, all=T)
```

Converting NA's to zero
```{r warning=FALSE, message=FALSE}
data_hu[is.na(data_hu)]=0
```

Formatting the data as required by deseq2 and converting NA's to zero
```{r warning=FALSE, message=FALSE}
data <- pivot_wider(data_hu, names_from = condition, values_from = nreads)
data[is.na(data)] <- 0
```

Converting data to matrix
```{r warning=FALSE, message=FALSE}
countData <- as.matrix(data[ ,-1])
colnames(countData)=c("low_rin_2", "low_rin_1", "high_rin_2", "high_rin_1")
rownames(countData)=data$Transcript
```

making a DGE object
```{r warning=FALSE, message=FALSE}
group <- c("low","low","high", "high")
d <- DGEList(counts=countData,group=factor(group))
```

Keeping the transcripts that have greater than 5 reads in at least 2 conditions
```{r warning=FALSE, message=FALSE}
dim(d)
d.full <- d
head(d.full$counts)
apply(d.full$counts, 2, sum)

keep <- rowSums(d.full$counts > 5) >= 2
d.full <- d.full[keep,]
dim(d.full)
```

Counts per million 
```{r warning=FALSE, message=FALSE}
cpm_data <- cpm(d.full)
```

Normalising the counts
```{r warning=FALSE, message=FALSE}
d.full <- calcNormFactors(d.full)
```

Exploring the data to see if the replicates are similar
```{r warning=FALSE, message=FALSE}
plotMDS(d.full, method = "bcv", col=as.numeric(d.full$samples$group))
```

Estimating common dispersion
```{r warning=FALSE, message=FALSE}
d1 <- estimateCommonDisp(d.full, verbose = T)
names(d1)
```

Estimating Tagwise dispersion
```{r warning=FALSE, message=FALSE}
d1 <- estimateTagwiseDisp(d1)
names(d1)
```

Plotting BCV- plots the tagwise biological coefficient of variation (square root of dispersions) against log2-CPM
```{r warning=FALSE, message=FALSE}
plotBCV(d1)
```

Differential expression-
Top tags extracts the most differentially expressed genes either ranked by the p-value or by absolute log-fold change
```{r warning=FALSE, message=FALSE}
et <- exactTest(d1)
results <- topTags(et, n = nrow(countData), sort.by = "none")
```

decideTestDGE identifies which genes are significantly differentially expressed from an edgeR fit object containing p-values and test statistics
total number od differentially expressed genes at FDR < 0.05
```{r warning=FALSE, message=FALSE}
de1 <- decideTestsDGE(et, adjust.method = "BH", p.value = 0.05, lfc = 1)
summary(de1)
```

The function plotsmear generates a plot of the tagwise fold change against log-CPM (analogous to MA plot). DE tags are highlighted on the plot
```{r warning=FALSE, message=FALSE}
deltags <- rownames(d1)[as.logical(de1)]
plotSmear(et, de.tags = deltags)
abline(h=c(-1,1), col="blue")
```


Without filtration
```{r warning=FALSE, message=FALSE}
res_without_nano <- edger_to_df(results)
colnames(res_without_nano)[1] <- "Transcript"  # FDR is adj p_value

res_without_nano$diffexpressed <- "NO"
res_without_nano$diffexpressed[res_without_nano$logFC > 1 & res_without_nano$FDR < 0.05] <- "UP"
res_without_nano$diffexpressed[res_without_nano$logFC < -1 & res_without_nano$FDR < 0.05] <- "DOWN"

z <- res_without_nano %>% 
  filter(diffexpressed=="UP")
```

With filtration
```{r warning=FALSE, message=FALSE}
res_without_nano_f <- edger_to_df(results)
colnames(res_without_nano_f)[1] <- "Transcript"  # FDR is adj p_value

res_without_nano_f$diffexpressed <- "NO"
res_without_nano_f$diffexpressed[res_without_nano_f$logFC > 1 & res_without_nano_f$FDR < 0.05] <- "UP"
res_without_nano_f$diffexpressed[res_without_nano_f$logFC < -1 & res_without_nano_f$FDR < 0.05] <- "DOWN"

z <- res_without_nano_f %>% 
  filter(diffexpressed=="UP")
```

#####################################################################################
Exploring the dets- found after filtration

res_without_nano_f
separating dets, up and down regulated transcripts
```{r warning=FALSE, message=FALSE}
z_up <- res_without_nano_f %>% 
  filter(diffexpressed=="UP")
z_down <- res_without_nano_f %>% 
  filter(diffexpressed=="DOWN")
z_det <- res_without_nano_f %>% 
  filter(diffexpressed=="UP" | diffexpressed=="DOWN")
```

changing the colname
```{r warning=FALSE, message=FALSE}
colnames(z_up)[1] <- "Transcript"
colnames(z_down)[1] <- "Transcript"
colnames(z_det)[1] <- "Transcript"
```

Merging datasets to get transcript length or read length
```{r warning=FALSE, message=FALSE}
merged_tx <- merge(x=udeg_rep1_txlen, y=z_det, by="Transcript")

merged_low <- merge(x=lrin_all_filtering_rep1, y=z_det, by="Transcript") #udeg_rep1 is data after filtration
merged_high <- merge(x=hrin_all_filtering_rep1, y=z_det, by="Transcript") #hdeg_rep1 is data after filtration
```

merging datasets to get read lengths - merged to the datasets after filtration
```{r warning=FALSE, message=FALSE}
up_merge_u <- merge(x=udeg_rep1_txlen, y=z_up, by="Transcript")
down_merge_u <- merge(x=udeg_rep1_txlen, y=z_down, by="Transcript")

up_merge_h <- merge(x=hdeg_rep1, y=z_up, by="Transcript")
down_merge_h <- merge(x=hdeg_rep1, y=z_down, by="Transcript")
```

transcript length distribution
```{r warning=FALSE, message=FALSE}
ggplot(up_merge, aes(x=Transcript_len)) + geom_histogram(binwidth = 1, col="blue") + ggtitle("Upregulated transcripts- TL distribution- undeg_rep 1") + theme_bw()+ theme(axis.text=element_text(size=14), axis.title=element_text(size=16))

ggplot(down_merge, aes(x=Transcript_len)) + geom_histogram(binwidth = 1, col="blue") + ggtitle("Down regulated transcripts- TL distribution- undeg_rep 1") + theme_bw()+ theme(axis.text=element_text(size=14), axis.title=element_text(size=16))

ggplot() + aes(Transcript_length) + geom_histogram(binwidth = 1, col="blue") + ggtitle("DETs- TL distribution- undeg_rep 1") + theme_bw()+ theme(axis.text=element_text(size=14), axis.title=element_text(size=16))
```

Read length distribution
```{r warning=FALSE, message=FALSE}
ggplot(merged_low, aes(x=new_ReadLen)) + geom_histogram(binwidth = 1, col="blue") + ggtitle("DETs- Distribution plot of read lengths- Low RIN") + theme_bw() + labs(x="Read Length", y="Number of Reads") + theme(axis.text=element_text(size=16), axis.title=element_text(size=18)) + theme(plot.title = element_text(size=18, face = "bold"))+ xlim(0,5000)

ggplot(merged_high, aes(x=new_ReadLen)) + geom_histogram(binwidth = 1, col="blue") + ggtitle("DETs- Distribution plot of read lengths- High RIN") + theme_bw() + labs(x="Read Length", y="Number of Reads") + theme(axis.text=element_text(size=16), axis.title=element_text(size=18)) + theme(plot.title = element_text(size=18, face = "bold")) + xlim(0,5000)

ggplot(hrin_all_filtering_rep1, aes(x=new_ReadLen)) + geom_histogram(binwidth = 1, col="blue") + ggtitle("Distribution plot of read lengths- High RIN") + theme_bw() + labs(x="Read Length", y="Number of Reads") + theme(axis.text=element_text(size=16), axis.title=element_text(size=18)) + theme(plot.title = element_text(size=18, face = "bold")) + xlim(0,5000)

ggplot(lrin_all_filtering_rep1, aes(x=new_ReadLen)) + geom_histogram(binwidth = 1, col="blue") + ggtitle("Distribution plot of read lengths- Low RIN") + theme_bw() + labs(x="Read Length", y="Number of Reads") + theme(axis.text=element_text(size=16), axis.title=element_text(size=18)) + theme(plot.title = element_text(size=18, face = "bold"))+ xlim(0,5000)
```

