---
title: "rat_cerebellum"
output: html_document
date: "2023-09-28"
editor_options: 
  chunk_output_type: console
---

loading libraries
```{r warning=FALSE, message=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(GenomicFeatures)
  library(rtracklayer)
  library(dplyr)
  library(Rsamtools)
  library(glmnet)
  library(ggpubr)
  library(visreg)
  library(DESeq2)
  library(apeglm)
  library(edgeR)
  library(coriell)
})
```

loading data
```{r warning=FALSE, message=FALSE}
b_rin1l <- set_1
b_rin2l <- set_1
b_rin1h <- set_1
b_rin2h <- set_1

a_rin1l <- lrin_all_filtering_rep1
a_rin2l <- lrin_all_filtering_rep2
a_rin1h <- hrin_all_filtering_rep1
a_rin2h <- hrin_all_filtering_rep2
```

Read length distribution plots

```{r warning=FALSE, message=FALSE}
ggplot(b_rin1h, aes(x=ReadLen)) + geom_histogram(binwidth = 1, col="blue") + ggtitle("Read Length Distribution- Rat Cerebellum- RIN 8.4") + theme_bw()+ theme(axis.text=element_text(size=16), axis.title=element_text(size=18)) + theme(plot.title = element_text(size=18, face = "bold")) + labs(x="Read Length", y="Number of Reads") + xlim(0,7500) + ylim(0,1750)

ggplot(b_rin1l, aes(x=ReadLen)) + geom_histogram(binwidth = 1, col="blue") + ggtitle("Read Length Distribution- Rat Cerebellum- RIN 7.0") + theme_bw()+ theme(axis.text=element_text(size=16), axis.title=element_text(size=18)) + theme(plot.title = element_text(size=18, face = "bold")) + labs(x="Read Length", y="Number of Reads") + ylim(0,1750) + xlim(0,7500)
```

Before filtering
Selecting required columns
```{r warning=FALSE, message=FALSE}
bl <- b_rin1l %>% 
  dplyr::select(Transcript, ReadLen)
bh <- b_rin1h %>% 
  dplyr::select(Transcript, ReadLen)

colnames(bl)[2] <- "ReadLen_l"
colnames(bh)[2] <- "ReadLen_h"

bl$rin <- "7.0"
bh$rin <- "8.4"

merged <- merge(x=bl, y=bh, all=T)

wilcox.test(merged$ReadLen~merged$rin, alternative="less")

ggplot(merged) + aes(x=rin, y=ReadLen) + geom_boxplot()

qqnorm(bl$ReadLen)

qqnorm(bh$ReadLen)
```

Ordering the Read len columns
```{r warning=FALSE, message=FALSE}
bl2 <- bl[order(bl$ReadLen), ]
bh2 <- bh[order(bh$ReadLen), ]
```

Normalising between 0 and 1
```{r warning=FALSE, message=FALSE}
bl2$percent=(1:nrow(bl2))/nrow(bl2)
bh2$percent=(1:nrow(bh2))/nrow(bh2)
```

Before filtering
```{r warning=FALSE, message=FALSE}
ggplot()+
  geom_point(data = bl2, aes(x=percent, y=ReadLen), color="green")+
  geom_point(data = bh2, aes(x=percent, y=ReadLen), color="blue") + theme_bw() + ggtitle("Rat cerebellum- before filtering") + ylim(0,30000)
```

After filtering
Selecting required columns
```{r warning=FALSE, message=FALSE}
al <- a_rin1l %>% 
  dplyr::select(Transcript, new_ReadLen)
ah <- a_rin1h %>% 
  dplyr::select(Transcript, new_ReadLen)
```

Ordering the Read len columns
```{r warning=FALSE, message=FALSE}
al2 <- al[order(al$new_ReadLen), ]
ah2 <- ah[order(ah$new_ReadLen), ]
```

Normalising between 0 and 1
```{r warning=FALSE, message=FALSE}
al2$percent=(1:nrow(al2))/nrow(al2)
ah2$percent=(1:nrow(ah2))/nrow(ah2)
```

Before filtering
```{r warning=FALSE, message=FALSE}
ggplot()+
  geom_point(data = al2, aes(x=percent, y=new_ReadLen), color="green")+
  geom_point(data = ah2, aes(x=percent, y=new_ReadLen), color="blue") + theme_bw() + ggtitle("Rat cerebellum- after filtering") + ylim(0,30000)
```



3' end distribution
Before filtering
Selecting required columns
```{r warning=FALSE, message=FALSE}
bl <- b_rin1l %>% 
  dplyr::select(Transcript, end_without_sc)
bh <- b_rin1h %>% 
  dplyr::select(Transcript, end_without_sc)
```

Ordering the Read len columns
```{r warning=FALSE, message=FALSE}
bl2 <- bl[order(bl$end_without_sc), ]
bh2 <- bh[order(bh$end_without_sc), ]
```

Normalising between 0 and 1
```{r warning=FALSE, message=FALSE}
bl2$percent=(1:nrow(bl2))/nrow(bl2)
bh2$percent=(1:nrow(bh2))/nrow(bh2)
```

Before filtering
```{r warning=FALSE, message=FALSE}
ggplot()+
  geom_point(data = bl2, aes(x=percent, y=end_without_sc), color="green")+
  geom_point(data = bh2, aes(x=percent, y=end_without_sc), color="blue") + theme_bw() + ggtitle("Rat cerebellum- before filtering")
```

After filtering
Selecting required columns
```{r warning=FALSE, message=FALSE}
al <- a_rin1l %>% 
  dplyr::select(Transcript, new_3_end)
ah <- a_rin1h %>% 
  dplyr::select(Transcript, new_3_end)
```

Ordering the Read len columns
```{r warning=FALSE, message=FALSE}
al2 <- al[order(al$new_3_end), ]
ah2 <- ah[order(ah$new_3_end), ]
```

Normalising between 0 and 1
```{r warning=FALSE, message=FALSE}
al2$percent=(1:nrow(al2))/nrow(al2)
ah2$percent=(1:nrow(ah2))/nrow(ah2)
```

After filtering
```{r warning=FALSE, message=FALSE}
ggplot()+
  geom_point(data = al2, aes(x=percent, y=new_3_end), color="green")+
  geom_point(data = ah2, aes(x=percent, y=new_3_end), color="blue") + theme_bw() + ggtitle("Rat cerebellum- after filtering") + ylim(0,32000)
```


