---
title: "3' end filtering_diff approach"
output: html_document
date: "2023-07-31"
editor_options: 
  chunk_output_type: console
---

INCOMPLETE- NOT COMPLETELY SOLVED YET
This script plots first the distribution plot of the 3' end of the reads of the transcript and then calculates the saturation point. 

Loading Libraries
```{r warning=FALSE, message=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(GenomicFeatures)
  library(rtracklayer)
  library(dplyr)
  library(Rsamtools)
  library(glmnet)
  library(ggpubr)
  library(visreg)
  library(DESeq2)
  library(apeglm)
  library(slider)
})
```

```{r warning=FALSE, message=FALSE}
hdeg_normal <- read.csv("/home/bhavika/Desktop/Nanograd/Data_2/heavily_degraded.csv")
hdeg_nano <- read.csv("/home/bhavika/Desktop/Nanograd/Data_2/primary_align_nano_hdeg_pass1.csv")

udeg_normal <- read.csv("/home/bhavika/Desktop/Nanograd/Data_2/undegraded.csv")
udeg_nano <- read.csv("/home/bhavika/Desktop/Nanograd/Data_2/primary_align_nano_udeg_pass1.csv")
```

Selecting required column
```{r warning=FALSE, message=FALSE}
test_1h <- hdeg_normal %>% 
  dplyr::select(Read, Transcript, ReadLen, end_without_sc)
test_1u <- udeg_normal %>% 
  dplyr::select(Read, Transcript, ReadLen, end_without_sc)
```

Filtering transcripts
```{r warning=FALSE, message=FALSE}
test_filter_h <- test_1h %>% 
  filter(Transcript=="ENST00000361575.4")
test_filter_u <- test_1u %>% 
  filter(Transcript=="ENST00000361575.4")
```

distribution plot of the 3' ends - histogram
```{r warning=FALSE, message=FALSE}
ggplot(test_filter_h, aes(x=end_without_sc)) + geom_histogram(binwidth = 1, col="blue") + labs(x="3' end without sc") + ggtitle("Hdeg-ENST00000361575.4") + theme_bw()

ggplot(test_filter_u, aes(x=end_without_sc)) + geom_histogram(binwidth = 1, col="blue") + labs(x="3' end without sc") + ggtitle("Udeg-ENST00000361575.4") + theme_bw()
```


Ordering the 3' end column and calculating the frequency of each 3' end
```{r warning=FALSE, message=FALSE}
test_order_h <- test_filter_h[order(test_filter_h$end_without_sc, decreasing = TRUE),]
test_order_u <- test_filter_u[order(test_filter_u$end_without_sc, decreasing = TRUE),]

test_2h <- test_order_h %>% 
  group_by(end_without_sc) %>% 
  mutate(number=n())
test_2u <- test_order_u %>% 
  group_by(end_without_sc) %>% 
  mutate(number=n())
```

Taking the cumulative sum of the frequency 
```{r warning=FALSE, message=FALSE}
test_2u$cum <- cumsum(test_2u$number)
p <- function(x,n){
  len.diff <- n - length(x)
  c(rep(NA, len.diff), x)
}
test_2u$dif <- p(diff(test_2u$cum, lag=20), length(test_2u$cum))
test_2u$dif[is.na(test_2u$dif)] <- 0

test_2u$saturation <- test_2u$end_without_sc[which.max(test_2u$dif)] + 20

```

