---
title: "soft_clip_plots"
output: html_document
date: "2023-04-27"
editor_options: 
  chunk_output_type: console
---

This script is for distribution plots of 5' and 3' end with and without soft clips for the full data set of undegraded and heavily degraded data. 

loading libraries
```{r warning=FALSE, message=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(GenomicFeatures)
  library(rtracklayer)
  library(dplyr)
  library(Rsamtools)
  library(glmnet)
  library(ggpubr)
  library(visreg)
  library(DESeq2)
  library(apeglm)
})
```

loading data
```{r warning=FALSE, message=FALSE}
heavily_degraded <- read.csv("/home/bhavika/Desktop/Nanograd/Data/csv_files/heavily_degraded.csv")
undegraded <- read.csv("/home/bhavika/Desktop/Nanograd/Data/csv_files/undegraded.csv")
mildly_degraded <- read.csv("/home/bhavika/Desktop/Nanograd/Data/csv_files/mildly_degraded.csv")
```

5' END

Dropping NA's (if any) from the columns
```{r warning=FALSE, message= FALSE}
test_h <- heavily_degraded %>% 
  drop_na(start_without_sc)
test_u <- undegraded %>% 
  drop_na(start_without_sc)
```

selecting transcripts for which plot has to be generated. 
```{r warning=FALSE, message=FALSE}
filter_h <- test_h %>% 
  filter(Transcript=="ENST00000525828.1")
filter_u <- test_u %>% 
  filter(Transcript=="ENST00000525828.1")
```

Distribution plot for 5' end with sc and mentioning the Transcript length (TL) on the plot 
```{r warning=FALSE, message=FALSE}
a <- ggplot(filter_u, aes(x=StartCoord)) + geom_histogram(aes(y=stat(density))) + geom_density(col="red") + ggtitle("Undegraded- 5' end with soft clip")+ scale_x_continuous(limits = c(-100, 700)) + geom_vline(xintercept = 664, colour="green") + labs(x="5' end with soft clip") + theme_bw()

a <- ggplot(filter_h, aes(x=StartCoord)) + geom_histogram(aes(y=stat(density))) + geom_density(col="red") + ggtitle("Heavily degraded- 5' end with soft clip") + scale_x_continuous(limits = c(-100, 700)) + geom_vline(xintercept = 664, colour="green") + labs(x="5' end with soft clip") + theme_bw()
```

Checking the binwidth of the histogram
```{r warning=FALSE, message=FALSE}
gg_build <- ggplot_build(a)
bin_width <- gg_build$data[[1]]$xmax - gg_build$data[[1]]$xmin
unique(bin_width)
```

Distribution plot for 5' end without sc and mentioning TL on the plot
```{r warning=FALSE, message=FALSE}
b <- ggplot(filter_u, aes(x=start_without_sc)) + geom_histogram(aes(y=stat(density))) + geom_density(col="red") + ggtitle("Undegraded- 5' end without soft clip") + scale_x_continuous(limits= c(-100, 700)) + geom_vline(xintercept = 664, colour="green") + labs(x="5' end without soft clip") + theme_bw()

b <- ggplot(filter_h, aes(x=start_without_sc)) + geom_histogram(aes(y=stat(density))) + geom_density(col="red") + ggtitle("Heavily degraded- 5' end without soft clip") + scale_x_continuous(limits= c(-100, 700)) +geom_vline(xintercept = 664, colour="green") + labs(x="5' end without soft clip") + theme_bw()
```

Checking the binwidth of histograms
```{r warning=FALSE, message=FALSE}
gg_build <- ggplot_build(b)
bin_width <- gg_build$data[[1]]$xmax - gg_build$data[[1]]$xmin
unique(bin_width)
```

3' END 

Dropping NA's (if any) from the columns

```{r warning=FALSE, message=FALSE}
test_h <- heavily_degraded %>% 
  drop_na(end_without_sc)
test_u <- undegraded %>% 
  drop_na(end_without_sc)
```

Selecting the transcripts for which plot has to be generated.
```{r warning=FALSE, message=FALSE}
filter_h <- test_h %>% 
  filter(Transcript=="ENST00000525828.1")
filter_u <- test_u %>% 
  filter(Transcript=="ENST00000525828.1")
```

Distribution plot for 3' with soft clip and mentioning TL on the plot
```{r warning=FALSE, message=FALSE}
c <- ggplot(filter_u, aes(x=end_with_sc)) + geom_histogram(aes(y=stat(density))) + geom_density(col="red") + ggtitle("Undegraded- 3' end with soft clip") + scale_x_continuous(limits = c(0,1000)) + geom_vline(xintercept = 664, colour="green") + labs(x="3' end with soft clip") + theme_bw()

c <- ggplot(filter_h, aes(x=end_with_sc)) + geom_histogram(aes(y=stat(density))) + geom_density(col="red") + ggtitle("Heavily degraded- 3' end with soft clip") + scale_x_continuous(limits = c(0,1000)) + geom_vline(xintercept = 664, colour="green") + labs(x="3' end with soft clip") + theme_bw()
```

Checking binwidth of the histogram
```{r warning=FALSE, message=FALSE}
gg_build <- ggplot_build(c)
bin_width <- gg_build$data[[1]]$xmax - gg_build$data[[1]]$xmin
unique(bin_width)
```

Distribution plots for 3' without soft clip and mentioning TL on the plot
```{r warning=FALSE, message=FALSE}
d <- ggplot(filter_u, aes(x=end_without_sc)) + geom_histogram(aes(y=stat(density))) + geom_density(col="red") + ggtitle("Undegraded- 3' end without soft clip") + scale_x_continuous(limits = c(0,1000)) + geom_vline(xintercept = 664, colour="green") + labs(x="3' end without soft clip") + theme_bw()

d <- ggplot(filter_h, aes(x=end_without_sc)) + geom_histogram(aes(y=stat(density))) + geom_density(col="red") + ggtitle("Heavily degraded- 3' end without soft clip") + scale_x_continuous(limits = c(0,1000)) + geom_vline(xintercept = 664, colour="green") + labs(x="3' end without soft clip") + theme_bw()
```

Checking binwidth of the histogram
```{r warning=FALSE, message=FALSE}
gg_build <- ggplot_build(d)
bin_width <- gg_build$data[[1]]$xmax - gg_build$data[[1]]$xmin
unique(bin_width)
```
