---
title: "unfiltered_filtered_plots"
output: html_document
date: "2023-05-12"
editor_options: 
  chunk_output_type: console
---

This script plots unfiltered and filtered read len plots of transcripts. The plot also includes mean, median and slope of the read lens. 

Formula for slope calculation used = (number of reads)/(max read len - min read len)

Loading Libraries
```{r warning=FALSE, message=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(GenomicFeatures)
  library(rtracklayer)
  library(dplyr)
  library(Rsamtools)
  library(glmnet)
  library(ggpubr)
  library(visreg)
  library(DESeq2)
  library(apeglm)
})
```

loading replicate 1 data
```{r warning=FALSE, message=FALSE}
hdeg_rep1 <- read.csv("/home/bhavika/Desktop/Nanograd/Data/csv_files/heavily_degraded.csv")
mdeg_rep1 <- read.csv("/home/bhavika/Desktop/Nanograd/Data/csv_files/mildly_degraded.csv")
udeg_rep1 <- read.csv("/home/bhavika/Desktop/Nanograd/Data/csv_files/undegraded.csv")
```

loading replicate 2 data
```{r warning=FALSE, message=FALSE}
udeg_rep_2 <- read.csv("/home/bhavika/Desktop/Nanograd/Data/csv_files/undegraded_rep_2.csv")
mdeg_rep_2 <- read.csv("/home/bhavika/Desktop/Nanograd/Data/csv_files/mildly_degraded_rep_2.csv")
hdeg_rep_2 <- read.csv("/home/bhavika/Desktop/Nanograd/Data/csv_files/heavily_degraded_rep_2.csv")
```

loading max lens data
```{r warning=FALSE, message=FALSE}
hdeg_max <- read.csv("/home/bhavika/Desktop/Nanograd/Data/csv_files/hdeg_max_lens.csv")
mdeg_max <- read.csv("/home/bhavika/Desktop/Nanograd/Data/csv_files/mdeg_max_lens.csv")
udeg_max <- read.csv("/home/bhavika/Desktop/Nanograd/Data/csv_files/udeg_max_lens.csv")
```

Filtering the data for selected transcript
```{r warning=FALSE, message=FALSE}
tx_h <- hdeg_rep1 %>% 
  filter(Transcript=="ENST00000361575.4")
tx_u <- udeg_rep1 %>% 
  filter(Transcript=="ENST00000361575.4")
```

Selecting required columns- Transcript and Readlen
```{r warning=FALSE, message=FALSE}
lh <- tx_h %>% 
  dplyr::select(Transcript, ReadLen)
lu <- tx_u %>% 
  dplyr::select(Transcript, ReadLen)
```

Ordering the read len column
```{r warning=FALSE, message=FALSE}
lh2 <- lh[order(lh$ReadLen), ]
lu2 <- lu[order(lu$ReadLen), ]
```

Normalising the reads between 0 and 1
```{r warning=FALSE, message=FALSE}
lh2$count=(1:nrow(lh2))/nrow(lh2)
lu2$count=(1:nrow(lu2))/nrow(lu2)
```

Unfiltered plot
```{r warning=FALSE, message=FALSE}
ggplot()+
  geom_point(data = lh2, aes(x=count, y=ReadLen), color="red")+ geom_hline(yintercept = 391, col="red") +
  geom_point(data = lu2, aes(x=count, y=ReadLen), color="blue") + geom_hline(yintercept = 393, col="blue") + theme_bw() + ggtitle("Unfiltered")
```

Calculating the mean, median and slope
```{r warning=FALSE, message=FALSE}
test_1 <- lh %>% 
  dplyr::group_by(Transcript) %>% 
  summarise(median=median(ReadLen))
test_1

test_2 <- lu %>% 
  dplyr::group_by(Transcript) %>% 
  summarise(median=median(ReadLen))
test_2

test_1 <- lh %>% 
  dplyr::group_by(Transcript) %>% 
  summarise(mean=mean(ReadLen))
test_1

test_2 <- lu %>% 
  dplyr::group_by(Transcript) %>% 
  summarise(mean=mean(ReadLen))
test_2
```

Adding Read len column in the data on max len dataset
```{r warning=FALSE, message=FALSE}
hdeg_max$ReadLen <- hdeg_max$MapLen + hdeg_max$DelLen + hdeg_max$InsertLen
udeg_max$ReadLen <- udeg_max$MapLen + udeg_max$DelLen + udeg_max$InsertLen
```

Adding map len and del len conditions
```{r warning=FALSE, message=FALSE}
hdeg_filter <- hdeg_max %>% 
  mutate(filter_1=ifelse(MapLen > 200 & MaxDelLen > 160 , "discard", "not discard"))

udeg_filter <- udeg_max %>% 
  mutate(filter_1=ifelse(MapLen > 200 & MaxDelLen > 160, "discard", "not discard"))
```

map len and insert len conditions
```{r warning=FALSE, message=FALSE}
hdeg_filter <- hdeg_max %>% 
  mutate(filter_1=ifelse(MapLen > 200 & MaxInsertLen > 60 , "discard", "not discard"))

udeg_filter <- udeg_max %>% 
  mutate(filter_1=ifelse(MapLen > 200 & MaxInsertLen > 60, "discard", "not discard"))
```

Filtering the data for selected transcript
```{r warning=FALSE, message=FALSE}
tx_h <- hdeg_filter %>% 
  filter(Transcript=="ENST00000216146.9") %>% 
  filter(filter_1=="not discard")

tx_u <- udeg_filter %>% 
  filter(Transcript=="ENST00000216146.9") %>% 
  filter(filter_1=="not discard")
```

Selecting required columns
```{r warning=FALSE, message=FALSE}
lh <- tx_h %>% 
  dplyr::select(Transcript, ReadLen)
lu <- tx_u %>% 
  dplyr::select(Transcript, ReadLen)
```

Ordering the Read len columns
```{r warning=FALSE, message=FALSE}
lh2 <- lh[order(lh$ReadLen), ]
lu2 <- lu[order(lu$ReadLen), ]
```

Normalising between 0 and 1
```{r warning=FALSE, message=FALSE}
lh2$count=(1:nrow(lh2))/nrow(lh2)
lu2$count=(1:nrow(lu2))/nrow(lu2)
```

Filtered plot of map len and max del len condition
```{r warning=FALSE, message=FALSE}
ggplot()+
  geom_point(data = lh2, aes(x=count, y=ReadLen), color="red")+
  geom_point(data = lu2, aes(x=count, y=ReadLen), color="blue") + theme_bw() + ggtitle("Filtered- condition (MapLen > 200 & MaxDelLen > 160)")
```

Filtered plot of map len and max insert len condition
```{r warning=FALSE, message=FALSE}
ggplot()+
  geom_point(data = lh2, aes(x=count, y=ReadLen), color="red")+
  geom_point(data = lu2, aes(x=count, y=ReadLen), color="blue") + theme_bw() + ggtitle("ENST00000216146.9- Filtered- condition (MapLen > 200 & MaxInsertLen > 60)") + ylim(0,1700)
```

Calculating the mean, median and slope
```{r warning=FALSE, message=FALSE}
test_1 <- lh %>% 
  dplyr::group_by(Transcript) %>% 
  summarise(mean=mean(ReadLen))
test_1

test_2 <- lu %>% 
  dplyr::group_by(Transcript) %>% 
  summarise(mean=mean(ReadLen))
test_2

test_1 <- lh %>% 
  dplyr::group_by(Transcript) %>% 
  summarise(median=median(ReadLen))
test_1

test_2 <- lu %>% 
  dplyr::group_by(Transcript) %>% 
  summarise(median=median(ReadLen))
test_2

max(lh$ReadLen)
min(lh$ReadLen)

max(lu$ReadLen)
min(lu$ReadLen)
```

Distribution plot of read lens of a particular transcript to check the saturation point
Selecting the transcript
```{r warning=FALSE, message=FALSE}
sample <- hdeg_max %>% 
  filter(Transcript=="ENST00000234875.9")
sample_1 <- udeg_max %>% 
  filter(Transcript=="ENST00000234875.9")
```

Distribution plot of Read lens
```{r warning=FALSE, message=FALSE}
a <- ggplot(sample, aes(x=ReadLen)) + geom_histogram(aes(y=stat(density)), binwidth = 1, fill="blue") + geom_density(col="red") + labs(x="Read Length") + ggtitle("Heavy degraded- ENST00000234875.9") + theme_bw() + ylim(0.00,0.05) + xlim(0,2500) + geom_vline(xintercept = 2061, col="green")

b <- ggplot(sample_1, aes(x=ReadLen)) + geom_histogram(aes(y=stat(density)), binwidth = 1, fill="blue") + geom_density(col="red") + labs(x="Read Length") + ggtitle("Undegraded- ENST00000234875.9") + theme_bw()+ ylim(0.00,0.05) + xlim(0,2500) + geom_vline(xintercept = 2061, col="green")
```

Arranging the plots in one frame
```{r warning=FALSE, message=FALSE}
plot <- ggarrange(b, a, ncol = 1, nrow = 2)

annotate_figure(plot, top = text_grob("Distribution plot of Read Lengths", color = "red", face = "bold", size = 14))
```


